---
title: "Day 2 Interactive Activity"
author: "Heather"
format: html
editor: visual

execute: 
  warning: false
  echo: true
---

# Day 2

## Making and Indexing Data in R

### Making Vectors

Character Vectors

```{r}
# Make a vector of NCEAS Dogs
dogs <- c("teddy", "khora", "waffle", "banjo", "tallie")

#  check the class of the vector
class(dogs)
typeof (dogs)
```

Multi variable vector

```{r}
my_breakfast <- c("waffles", 2, TRUE, "coffee")
class(my_breakfast)
my_breakfast

```

Numeric vs Integer Vectors

```{r}
weights <- c(50, 60, 25, 35, 55)
class(weights)

weights_Integer <- c(50L, 60L, 25L, 35L, 55L)
class(weights_Integer)
```

You can use the is.\_\_\_\_\_\_\_ function to determine how the data is being stored by R

```{r}
is.numeric(weights_Integer)
```

Similarly, you can use the as.\_\_\_\_\_\_\_\_ function to redefine a set of data as a different kind of data. For example, R will read Zip Codes as numeric inputs. However, we know that Zip Codes are a category or a character data input.

```{r}
zipcodes <- c(93105, 93514, 93101, 93401)
class(zipcodes)

zipcodes_char <- as.character(zipcodes)
class(zipcodes_char)

```

### Create a vector with named elements

```{r}
dog_food <- c(teaddy = "purina", khora = "alpo", waffle = "fancy feast", banjo = "blue diamond",tallie = "farmer's dog")

class(dog_food)

portion_size <- c(teddy = "6", khora = "7", waffle = "9", banjo = "10", tallie = "15")
portion_size



```

### Call an element from a vector

```{r}
dog_food[2]
# R uses a 1-index, so the first element of the vector is denoted 1

# If an element is named, you can call the name as well
dog_food["khora"]
```

### Access ranges of values in a vector:

Use : to refer to a range (low:high)

```{r}
trees <- c("ponderosa", "lodgepole", "jeffery", "sugar", "bristlecone")
trees[3:5]
```

## FOR LOOP WARM-UP

```{r}
i <- seq(1:3)
trees[i]

```

```{r}
trees[3] <- "OAK"

trees
trees[i]
```

## Morning Activities: Part 2

Matrix Indexing

```{r}
fish_size <- matrix(c(0.8, 1.2, 0.4, 0.9), ncol=2, nrow=2, byrow = FALSE)
fish_size

class(fish_size)
str(fish_size)
typeof(fish_size)
```

```{r}
# Identify an element
fish_size[2,2]

# Identify a column
fish_size[,2]


```

An Edit!

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_LUNCH BREAK\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
